Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-09-22 Prelim/prelim.Rmd
worked on prelim slides

---
title: "Prelim"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

@@

- File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B}$. We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.

- $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors

- $Z_j$

@@

# Model

@@

Demonstration

# Managing large data

@@

@@

@@

- $m_{fl} = $

______________________________________

>>>>

---
title: "Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)"
author: "Brian Kundinger"
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex  
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, out.height="80%", out.width="80%", fig.align = 'center')
```

@@

```{r out.width="90%"}
knitr::include_graphics("fs_graphic.png")
```

@@

- File A with records indexed $i \in \{1, \ldots, n_A\}$ and file B with records $j \in \{1, \ldots, n_B\}$. We use $F$ features for linkage, with $L_f$ possible levels of agreement on feature $f$.

- $\Gamma \in \mathbb{R}^{n_A n_B \times F}$ matrix of comparison vectors where $\gamma_{ij}^f \in \{1, \ldots, L_f\}$

- $Z_j = \begin{cases} 
	i,  & \text{if records } i\in A \text{ and } j\in B \text{ refer to the same entity}; \\
	n_A + 1,  & \text{if record } j\in B \text{ does not have a match in file } A; \\
	\end{cases}$

@@

- Fellegi and Sunter (1969)
  - independent classification of all $n_A n_B$ record pairs. Transitive closure acheived through postprocessing

- Enamorado et al (2019): \texttt{fastlink} 
  - provides efficient and scalable FS model

- Sadinle (2017) - Beta Record Linkage (\texttt{BRL}) 
  - Bayesian method for bipartite matching, strictly enforces one-to-one matching. Does not scale well to large linkage tasks

# Fast Beta Linkage (fabl)

$$P(\Gamma|\mathbf{Z}, \mathbf{m}, \mathbf{u}) = \prod_{j=1}^{n_B}  \prod_{i=1}^{n_A}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)}$$

$$\mathbf{m_{f}} \sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{fL_f})$$
$$\mathbf{u_{f}} \sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{fL_f})$$
$$Z_j | \lambda =
\begin{cases} 
	\frac{1}{n_A}\lambda  & z_j \leq n_A; \\
	1-\lambda &  z_j  = n_A + 1 \\
\end{cases}$$
$$\lambda \sim \text{Beta}(\alpha_{\lambda}, \beta_{\lambda})$$

- Model specification allows for **parallel/distributed** computing, **hashing** of comparison vectors, and **storage efficient indexing (SEI)**

@@

- Recognize there are at most $P = \prod_{f = 1}^F L_f$ unique agreement patterns. When $(i,j)$ pair exhibits agreement pattern $p$, say $(i, j) \in h_p$

- Reduce data to sufficient statistics
  - $H_p = \sum_{i,j} \mathbf{1}_{(i, j) \in h_p}$
  - $H_{p_j} = \sum_{i} \mathbf{1}_{(i, j) \in h_p}$

- Gibbs sampler over each record pair has complexity $O(n_A \times n_B \times F)$, but sampler over agreement pattern has complexity $O(P \times n_B \times F)$

- Speeds up posterior calculations and sampling

# Managing Large Data

- **Distributed Computing** - Partition data in to chunks $\{A_I\}$ and $\{B_J\}$. Compare records, hash results, compute summary statistics in parallel, and synthesize results

- **Storage Efficient Indexing (SEI)** - Remove highly unlikely record labels from memory. Proper weights for calculations maintained through summary statistics $\{H_p\}$ and $\{H_{p_j}\}$

@@

```{r}
knitr::include_graphics("../notes/figures/sadinle_sim_plot.png")
```

@@

```{r}
knitr::include_graphics("../notes/figures/sadinle_speed_plot2.png")
```

# Simulation Results 3

```{r}
knitr::include_graphics("../notes/figures/speed_plot_fixed_nB.png")
```

@@

- Linkage in the face of duplicates within datasets

- Models that allow reliability of information to differ by subgroup in the data

- Linkage over blocked data (allows for much larger linkage tasks)

# Questions?

# Additional Slides

### 2021-09-22 methods_paper/fabl-paper.tex
worked on prelim slides

%% Reworking introduction for Brian's paper.

Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen2012data}. Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. In the literature, bipartite record linkage is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.

task consists of merging two disparate datafiles containing information on two overlapping sets of entities.

\begin{enumerate}
\item Let's re-visit the case study of Sadinle. Lay this out and the importance of it. 
\item Next, let's briefly review the landscape of the literature. 
\item State Brian's contributions. 
\item Then write a section on related work. 
\item Then write a section regarding the layout of the paper.
\end{enumerate}

\newpage

In many cases, this identification is itself the goal of an analysis, as in estimating human rights atrocities from disparate data sources, or linking voter information across election years. In others, we seek to perform some other statistical procedure, such as survival analysis or regression, using linked datasets, and one must  propagate linkage uncertainty into the subsequent analysis.

@@

Wherever possible, we reserve superscripts for denoting field and level,
while reserving subscripts for record indices. For example,
\(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the probability
distribution governing field \(f\) for matching records, and
\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
is product of the relevant of the appropriate \(\mathbf{m}\) parameters
for record pair \((i,j)\). We hope that these conventions avoid
overloaded notation in the likelihood and subsequent derivations.

>>>>

Record linkage is the task of identifying
duplicate records across multiple data sources. In many cases, this identification is itself the goal of an analysis, as in estimating human rights atrocities from disparate data sources, or linking voter information across election years. In others, we seek to perform some other statistical procedure, such as survival analysis or regression, using linked datasets, and one must  propagate linkage uncertainty into the subsequent analysis.

@@

%Wherever possible, we reserve superscripts for denoting field and level,
%while reserving subscripts for record indices. For example,
%\(\mathbf{m}^f = (m^{f1}, \ldots, m^{fL_f})\) is the probability
%distribution governing field \(f\) for matching records, and
%\(\mathbf{m}_{ij}= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(m^{fl}\right)^{\mathbf{1}_{\gamma_{ij}^f = l}} = P(\boldsymbol{\gamma}_{ij}|Z_j = i)\)
%is product of the relevant of the appropriate \(\mathbf{m}\) parameters
%for record pair \((i,j)\). We hope that these conventions avoid
%overloaded notation in the likelihood and subsequent derivations.
